\Large \bf Cyclone: Warp Speed Replication for Key Value Stores

### 2017-04-18 paper/paper.tex
updates to paper.

The final dimension we evaluate is of ganged operations. The primary purpose of
ganged operations is to avoid the need for distributed transactions to
manipulate what is a single shared memory image and therefore we were most
concerned about unloaded latency given the complexity of synchronizing different
replication quorums as well executing our rendezvous protocol on multiple
cores. We therefore setup an experiment where a single client reflecting the
unloaded case made ganged requests to the replicas. We varied the number of
cores participating in the request from one to the full complement of 32
application cores. Figure~\ref{fig:ganged} shows the results both using a single
level log as well as a two level log. The primary takeaway is that unloaded
latency increases slowly as we increase the number of active cores - to around
40 us from the baseline of 20 us. There are two causes of this. First, there is
a rapid increase from 1 to 8 active cores, the reason being that this
corresponds to an increase in the number of quorums that must be synchronized
on. There is some amount of synchronization involved in the userspace DPDK
driver for access to common resources for accessing the NIC and this affects the
ability to simultaneously issues replication messages on all quorums. The
smaller contribution to increasing latency that continues past the 8 core case
is due to cacheline pingponging when executing the rendevous. Both these sources
of latency drift could be corrected using replication quorums mapped to
dedicated NICs and more scalable rendesvous designs (such as with machine aware
broadcast trees~\cite{broadcast_tree}). However we deemed the complexity of such
optimizations unnecessary. The added latency for most cases is well under the
extra round trip delay, which would the minimum needed for a solution using
distributed transactions via two phase commit.

@@

>>>>

The final dimension we evaluate is of ganged operations. The primary purpose of
ganged operations is to avoid the need for distributed transactions to
manipulate what is a single shared memory image and therefore we were most
concerned about unloaded latency given the complexity of synchronizing different
replication quorums as well executing our rendezvous protocol on multiple
cores. We therefore setup an experiment where a single client reflecting the
unloaded case made ganged requests to the replicas. We varied the number of
cores participating in the request from one to the full complement of 32
application cores. Figure~\ref{fig:ganged} shows the results both using a single
level log as well as a two level log. The primary takeaway is that unloaded
latency increases slowly as we increase the number of active cores - to around
40 us from the baseline of 20 us. There are two causes of this. First, there is
a rapid increase from 1 to 8 active cores, the reason being that this
corresponds to an increase in the number of quorums that must be synchronized
on. There is some amount of synchronization involved in the userspace DPDK
driver for access to common resources for accessing the NIC and this affects the
ability to simultaneously issues replication messages on all quorums. The
smaller contribution to increasing latency that continues past the 8 core case
is due to cacheline pingponging when executing the rendevous. Both these sources
of latency drift could be corrected using replication quorums mapped to
dedicated NICs and more scalable rendezvous designs (such as with machine aware
broadcast trees~\cite{broadcast_tree}). However we deemed the complexity of such
optimizations unnecessary. The added latency for most cases is well under the
extra round trip delay, which would the minimum needed for a solution using
distributed transactions via two phase commit.

@@

We now evaluate Cyclone integrated with the Rocksdb persistent key value
store. Rocksdb is a complex industrial strength persistent key value store and
this means that it is accompanied by a complex array of performance tuning knobs
to get the best performance from flash. We were also aware that SSD performance
is slated for dramatic increases in the coming years with the introduction of
new memory technology such as 3DXPoint based Optane SSDs. To demonstrate that
Cyclone is future proof we eliminated flash media performance from the picture
as far as the key value store is concerned by placing all files for the key
value store (SSTables) on a RAMdisk, which presumably represents the limit in
performance for flash in the near future. All log files on secondary storage
however - both Rocksdb's own write ahead log and the alternative of Cyclone's
second level flashlog - are placed on the SSDs.

We evaluate two different request sizes: 8 byte keys with 8 byte values and 8
bytes keys with 256 byte values. We run 8 physical logs on 8 cores and devote 32
cores to making requests to Rocksdb. Since we are interested in performance of
the log, which is only used for update requests, our workload is 100\% writes.

Before evaluating with Rocksdb we measure the baseline performance of
replicating the log with Cycloine for the given request sizes - Rocksdb performs
a no-op. We note that in addition to the key and value, we are also logging
RocksDB specific request data such as operation type and the request header.
Figure~\ref{fig:kv_baseline} shows the baseline performance for the chosen
request sizes. With the smaller request size, Cyclone can conservatively sustain
close to a million requests a second at a latency of just under 25us. With the
larger request size, Cyclone can sustain around 350K requests a second, again at
a latency of just under 25 us. Armed with these baseline numbers we now examine
how well Cyclone performs with Rocksdb.

\begin{figure}
\includegraphics[scale=0.6]{results2/kv_baseline.pdf}
\caption{Baseline KV replication performance}
\label{fig:kv_baseline}
\end{figure}

The performance of Rocksdb with Cyclone for the small update workload is shown
in Figure~\ref{fig:rocksdb}. We consider four different settings. The line
labeled Rocksdb is the key value store running with no logging whatsover - a
system crash would lead to data loss. The line labeled Rocksdb/WAL is for
Rocksdb running with its write ahead logging turned on. The large gap between
these two is the overhead of the existing Rocksdb WAL solution. The line labeled
Rocksdb/Cyclone 1 way is a two level Cyclone log but without any
replication. The line almost exactly tracks the performance of Rockdb. As
suggested by the baseline replication performance, Cyclone is able to provide a
write ahead log with no overhead to Rocksdb. The line labeled Rocksdb/Cyclone 3
way is with 3-way replication turned on. Other than a 20us delta due to the
extra network round trip, the line almost exactly tracks the Rocksdb no logging
case. Cyclone therefore provides high availability to Rocksdb at a fraction of
the cost of its existing single machine write ahead log. We also repeat the
experiment for the larger update size in Figure~\ref{fig:rocksdb_256}. The
conclusions are identical: Cyclone solves Rocksdb's write ahead logging
problem.

\begin{figure}
\includegraphics[scale=0.6]{results2/rocksdb.pdf}
\caption{Rocksdb - small updates}
\label{fig:rocksdb}
\end{figure}

\begin{figure}
\includegraphics[scale=0.6]{results2/rocksdb_256.pdf}
\caption{Rocksdb - large updates}
\label{fig:rocksdb_256}
\end{figure}

Next, we consider the problem of supporting Rocksdb's write-batch operation that
atomically writes a set of key-value pairs into the KV store. In order to
perform the operation with Cyclone managing the log, the client must issue a
ganged operation across cores owning the keys in the write batch. This is in
contrast to baseline Rocksdb where the entire write batch can be sent to any
core. A key concern here was whether Cyclone would add any latency to the
operation due to the extra synchronization needed across replication quorums and
participating application cores. We examine the problem for the unloaded case
and small updates in Figure~\ref{fig:rocksdb_multi} for a single client with
increasing number of keys in the batch - till 32 keys that covers all
application cores. The line labeled Rocksdb is with no write-ahead logging. We
note an increasing latency for this baseline indicating Rocksdb takes longer
with larger key batches. For the other options - Rocksdb/wal has considerably
larger latency. Cyclone does an effective job of cutting down on this latency
even as it needs to pay a price for synchronizing multiple quorums and
application cores making it somewhat slower than running Rocksdb with no logging
for batched writes.

\begin{figure}
\includegraphics[scale=0.6]{results2/rocksdb_multi.pdf}
\caption{Rocksdb - batched writes}
\label{fig:rocksdb_multi}
\end{figure}

Finally we showcase the benefit of using Cyclone beyond pure performance as
compared to the existing single machine Rocksdb write ahead log. Cyclone brings
multi-machine availability with the ability to automatically failover. We
demonstrate this in Figure ~\ref{fig:timeline} that shows the timeline of a run
where we kill the server process on the leader replica. Cyclone is configured to
with a 30ms failure detection timeout after which the client library tries
replicas in turn to locate the leader - in this case it fails over in about
60ms.

\begin{figure}
\includegraphics[scale=0.6]{results2/failover.pdf}
\caption{Rocksdb - failover}
\label{fig:timeline}
\end{figure}
